---
title: "Bivens Burn Data Analysis"
author: "W.W. Dillon"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

Target Journal:

PNAS, Proc. B, Nature Communications

Neobiota comments: serotiny of long-leaf, fire-effects at different times, fire size~intensity, total diff in number of living trees in the plots at the end (pre-burn/post-burn)

Relationship between height and survival - tree level analysis


```{r setup, message=FALSE, echo=F}

knitr::opts_chunk$set(echo = F, warning = F, message = F)
knitr::opts_chunk$set(fig.width = 7.5, fig.height = 7.5)

```


```{r load packages, message=FALSE}

library(plyr)
library(dplyr)
library(tidyr)
library(readxl)
library(readr)
library(ggplot2)
library(viridis)

```

```{r load data, message=FALSE}

temperatures <- read_csv("data/temperatureID_data_noNA.csv")
treatments <- read_excel("data/Bivens.treatments.xlsx")
flame_heights <- read_excel("data/FlameHts.xlsx")
fuel_data <- read_excel("data/BivensBurnData_fuel.xlsx")
tree_survival <- read_csv("data/LLP_Surv_2018_corrected.csv")

```

```{r global values}

figure_width <-  7.5
figure_height <- 7.5

# define color ####
# invasion_color <- scale_color_viridis(begin = .15, end = .9, option = "E", discrete = T)
# invasion_fill <- scale_fill_viridis(begin = .15, end = .9, option = "E", discrete = T)

invasion_color <- scale_color_manual(values = c("deepskyblue","red"))
invasion_fill <- scale_fill_manual(values = c("deepskyblue","red"))

# define theme ####
def_theme <- theme(legend.title = element_blank(),
            legend.text = element_text(size = 14),
            # legend.position = "none",
            axis.text = element_text(size = 18),
            axis.title = element_text(size = 18),
            plot.title = element_text(size = 20),
            strip.background = element_blank(),
            strip.text = element_text(size = 14),
            panel.grid = element_blank()
            )

dodge <- position_dodge(width=0.9)

```


```{r data munging}

# plot treatments ####
treatments <- treatments %>% 
  mutate(water_trt = substr(Treatment, 1,1),
         veg_trt = substr(Treatment,2,2),
         water_trt = if_else(water_trt=="D", "drought", "ambient"),
         water_trt = factor(water_trt, levels = c("ambient","drought")),
         veg_trt = if_else(veg_trt=="C", "invaded", "reference"),
         veg_trt = factor(veg_trt, levels = c("reference","invaded"))
         )


# fuel load data ####
## 3 measurements at 3 locations (flame/temp sensor) in each plot
## average the 3 measurements for each location
fuel_height <- fuel_data %>% 
  select(Plot, Location, FH1:FH3) %>% 
  gather(measure, fuel_height_cm, FH1, FH2, FH3) %>% 
  mutate(measure = substr(measure, 3,3)) %>% 
  group_by(Plot, Location) %>% 
  summarise(avg_fuel_ht_cm = mean(fuel_height_cm, na.rm = TRUE)) %>% 
  ungroup(.)

litter_depth <- fuel_data %>% 
  select(Plot, Location, LD1:LD3) %>% 
  gather(measure, litter_depth_cm, LD1, LD2, LD3) %>% 
  mutate(measure = substr(measure, 3,3)) %>% 
  group_by(Plot, Location) %>% 
  summarise(avg_litter_depth_cm = mean(litter_depth_cm, na.rm = TRUE)) %>% 
  ungroup(.)

fuel_structure <- left_join(
  fuel_height, 
  litter_depth, 
  by = c("Plot","Location")
)

fuel_structure <- left_join(
  fuel_structure,
  treatments %>% select(-Treatment),
  by = "Plot"
)

## 3 measurements for each plot (3 flame/temp sensor locations)
fuel_mass <- fuel_data %>% 
  select(Plot, Location, freshwt, drywt) %>% 
  mutate(fuel_moisture = 100*(freshwt-drywt)/freshwt)

fuel_mass <- left_join(
  fuel_mass,
  treatments %>% select(-Treatment),
  by = "Plot"
)

fuel_measures <- left_join(
  fuel_mass,
  fuel_structure
)


# flame height data ####
## 2 flame heights for 3 locations in each plot 
flame_hts <- flame_heights %>% 
  # mutate(avg_flame_ht = rowMeans(cbind(Flame_ht1,Flame_ht2), na.rm = T)) %>% 
  gather(flameID, flameht_cm, Flame_ht1, Flame_ht2) %>% 
  group_by(Plot, Location) %>% 
  summarise(avg_flameht_cm = mean(flameht_cm)) %>% 
  ungroup(.)

flame_hts <- left_join(
  flame_hts, treatments, by = "Plot"
)
summary(flame_hts)

# skimr::skim(flame_hts %>% 
#               group_by(water_trt, veg_trt) %>% 
#               select(flameht_cm))


flame_heights_trt <- flame_hts %>%
  group_by(veg_trt, water_trt) %>% 
  summarise(
            nobs = length(avg_flameht_cm),
            se_flameht = sd(avg_flameht_cm)/sqrt(nobs),
            avg_flameht_cm = mean(avg_flameht_cm))

 # mean(c(72,82,233,272,246,195))
 # sd(c(72,82,233,272,246,195))

# dodge_width <- 0.8

fuel_flame_data <- left_join(
  fuel_measures,
  flame_hts)

# summarize temperature data ####

temperatures <- temperatures %>% 
  rename(Plot = plot_num, 
         Box_ID = box_number, 
         Location = location) %>% 
  mutate(Box_ID = as.numeric(substr(Box_ID, start = 4, stop = nchar(Box_ID))),
         probe_ht = factor(probe_ht, levels = c("50cm","25cm","0cm")))


temps_abv_100_data <- temperatures %>% 
  group_by(Plot, Location, probe_ht) %>% 
  filter(probe_tempC >= 100) %>% 
  summarise(seconds_abv_100 = length(probe_tempC),
            heat_flux_abv_100 = sum(probe_tempC)
            ) %>% 
  ungroup(.)

# temps_abv_100_data <- left_join(
#   temps_abv_100_data,
#   fuel_flame_data
# )


# temps_abv_100_data %>% 
#   group_by(water_trt, veg_trt, probe_ht) %>% 
#   select(seconds_abv_100) %>% 
#   skimr::skim(.)

max_temperatures <- temperatures %>% 
  group_by(Plot, Location, probe_ht) %>% 
  arrange(desc(probe_tempC)) %>% 
  summarise(max_temp = max(probe_tempC),
            max_temp_avg5 = mean(probe_tempC[1:5])) %>% 
  ungroup(.)

# max_temperatures <- left_join(
#   max_temperatures,
#   fuel_flame_data
)

# max_temperatures %>% 
#   group_by(water_trt, veg_trt, probe_ht) %>% 
#   select(max_temp) %>% 
#   skimr::skim(.)

temperatures <- left_join(
  temperatures,
  treatments,
  by = "Plot"
)

# tree damage/survival ####

# tree_survival <- tree_survival %>% 
#   filter(!is.na(pct_brown)) %>% 
#   select(-Ind, -Row, -Column)
# summary(tree_survival)
# 
# tree_survival %>% filter(is.na(diam2_mm))
# unique(tree_survival$dbh_cm)
# summary(tree_survival)

tree_dmg <- tree_survival %>% 
  filter(!is.na(total_limbs)) %>% 
  select(Plot, Ind., pct_brown, dead_limbs, total_limbs, char_ht_cm, ht2_cm) %>% 
  mutate(tree_id = seq_len(nrow(.)),
         pct_dead_limbs = 100*dead_limbs/total_limbs)
## Excluded dbh because it is missing for 29 of the trees

tree_dmg <- left_join(
  tree_dmg,
  select(treatments, -Treatment)
)

tree_dmg <- tree_dmg %>% 
  mutate(alive = if_else(pct_dead_limbs==100, 0, 1),
         dead = if_else(pct_dead_limbs==100, 1, 0),
         survival = if_else(pct_dead_limbs==100, 0, 1))

# tree_dmg %>% 
#   group_by(veg_trt, water_trt) %>% 
#   summarise(plot_totals = n_distinct(Plot))

# summary(tree_dmg)



# aggregate data ####

agg_data <- left_join(
  max_temperatures,
  temps_abv_100_data
  )
# summary(agg_data)
agg_data[is.na(agg_data)] <- 0

# agg_data <- left_join(agg_data, fuel_flame_data)

# summarize to plot-level

agg_max_temperatures <- max_temperatures %>% 
  group_by(Plot, probe_ht) %>% 
  summarise(avg_max_temp = mean(max_temp_avg5)) %>% 
  ungroup(.)

agg_temps_abv_100_data <- temps_abv_100_data %>% 
  mutate(seconds_abv_100 = ifelse(is.na(seconds_abv_100), 0, seconds_abv_100),
         heat_flux_abv_100 = ifelse(is.na(heat_flux_abv_100), 0, heat_flux_abv_100)) %>% 
  group_by(Plot, probe_ht) %>% 
  summarise(avg_s_abv_100 = mean(seconds_abv_100),
            avg_heat_flux100 = mean(heat_flux_abv_100)) %>% 
  ungroup(.)
# summary(agg_temps_abv_100_data)

agg_fuel_flame_data <- fuel_flame_data %>% 
  group_by(Plot, water_trt, veg_trt, Treatment) %>% 
  summarise(avg_drywt = mean(drywt),
            avg_fuel_moisture = mean(fuel_moisture),
            avg_fuel_ht_cm = mean(avg_fuel_ht_cm),
            avg_litter_depth_cm = mean(avg_litter_depth_cm),
            avg_flameht_cm = mean(avg_flameht_cm)) %>% 
  ungroup(.)


agg_plot_data <- left_join(
  agg_max_temperatures, agg_temps_abv_100_data
) %>% 
  left_join(., agg_fuel_flame_data)
summary(agg_plot_data)
# summary(filter(agg_plot_data, probe_ht=="50cm"))

tree_agg_plot_data <- left_join(
  tree_dmg,
  agg_plot_data
  )
# summary(tree_agg_plot_data)
# n_distinct(filter(tree_agg_plot_data, probe_ht=="50cm")$tree_id)
# summary(filter(tree_agg_plot_data, probe_ht=="0cm"))

tree_dmg %>% 
  group_by(Plot, water_trt, veg_trt) %>% 
  summarise(trees_alive = sum(alive, na.rm = TRUE),
            trees_dead = sum(dead, na.rm = TRUE),
            avg_tree_ht_cm = mean(ht2_cm, na.rm = TRUE))

agg_plot_data <- left_join(
  agg_plot_data,
  tree_dmg %>% 
  group_by(Plot, water_trt, veg_trt) %>% 
  summarise(trees_alive = sum(alive, na.rm = TRUE),
            trees_dead = sum(dead, na.rm = TRUE),
            avg_tree_ht_cm = mean(ht2_cm, na.rm = TRUE))
)
# summary(agg_plot_data)
# filter(agg_plot_data, is.na(trees_dead))$Plot
## 8 plots with no living trees at time of fires


```

```{r export data for science fieldtrip, eval=F}

fuel_measures %>%
  rename(dry_weight_g = avg_drywt,
         veg_moisture_pct = avg_moisture,
         veg_ht_cm = avg_fuel_ht_cm) %>% 
  write_csv(., "~/CodingEnvSci_UF_FieldTrip/coding_curriculum/data/bivens_veg_data_example.csv")

```

# Fuel condition summary

## Fuel load & moisture figures

```{r skim fuel load data, eval=FALSE}

# fuel_ht
# fuel_mass
fuel_mass %>% 
  group_by(water_trt, veg_trt) %>% 
  select(drywt, fuel_moisture) %>%
  skimr::skim(.)

```


```{r fuel load barchart, eval=FALSE}

(fuel_load_fig <- ggplot(
  fuel_mass %>% 
    group_by(water_trt, veg_trt) %>% 
    summarise(avg_drywt = mean(drywt*16),
              nobs = length(drywt),
              se_drywt = sd(drywt*16)/sqrt(nobs)), 
  aes(water_trt, avg_drywt, fill = veg_trt)) +
  invasion_color + 
  invasion_fill +
  geom_col(position = dodge) +
  geom_errorbar(aes(ymax = avg_drywt + se_drywt, ymin = avg_drywt - se_drywt), position = dodge, width = 0) +
  # geom_point(aes(color = veg_trt)) +
  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2)) +
  # facet_grid(probe_ht~.) +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab(expression(paste("Fuel load (g ", {m}^-2, ")"), sep = "")) +
  theme(legend.position = "top")
)
ggsave("figures/fuel_load_bar.png", height = 4, width = 5, dpi = 300)

```

```{r fuel load mean se, eval=FALSE}

ggplot(fuel_mass, aes(water_trt, drywt*16, fill = veg_trt)) +
  invasion_color + 
  invasion_fill +
  # geom_point(aes(color = veg_trt), position = "jitter", alpha = .3) +
  # geom_col(position = dodge) +
  # geom_errorbar(aes(ymax = avg_drywt + se_drywt, ymin = avg_drywt - se_drywt), position = dodge, width = 0) +
  # geom_point(aes(color = veg_trt)) +
  stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
  # scale_y_continuous(limits = c(0,190)) +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab(expression(paste("Fuel load (g ", {m}^-2, ")"), sep = "")) +
  theme(legend.position = "top")

```

Fuel loads are the average of three 25 cm^2^ samples per plot.
```{r fuel load boxplot}

## Location-level aggregation n=120, 30 per treatment
# ggplot(fuel_mass, aes(water_trt, .001*drywt*16)) +
#   # geom_violin(draw_quantiles = c(.25, .5, .75)) +
#   geom_boxplot(aes(color = veg_trt), notch = F, outlier.color = "white") +
#   geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .2)) +
#   invasion_fill +
#   invasion_color+
#   theme_bw() +
#   def_theme +
#   xlab("") +
#   ylab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
#   theme(legend.position = "top")

## Plot-level aggregation n=40, 10 per treatment
ggplot(filter(agg_plot_data, probe_ht=="0cm"), aes(water_trt, .001*avg_dry_wt*16)) +
  geom_boxplot(aes(color = veg_trt), notch = F, outlier.color = "white") +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .2)) +
  invasion_fill +
  invasion_color+
  theme_bw() +
  def_theme +
  xlab("") +
  ylab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
  theme(legend.position = "top")

```

```{r fuel load table}

knitr::kable(agg_plot_data %>% 
    group_by(water_trt, veg_trt) %>% 
    summarise(avg_drywt = mean(.001*avg_dry_wt*16),
              nobs = length(avg_dry_wt),
              se_drywt = sd(.001*avg_dry_wt*16)/sqrt(nobs)) %>% 
      select(-nobs),
    digits = 2, caption = "Average fuel loads by treatment from sample dry weights (kg/m2)")

```

```{r fuel moisture barchart, eval=FALSE}

(fuel_moisture_fig <- ggplot(
  fuel_mass %>% 
    group_by(water_trt, veg_trt) %>% 
    summarise(avg_moisture = mean(fuel_moisture),
              nobs = length(fuel_moisture),
              se_moisture = sd(fuel_moisture)/sqrt(nobs)),
  aes(water_trt, avg_moisture, fill = veg_trt)) +
  invasion_color + 
  invasion_fill +
  geom_col(position = dodge) +
  geom_errorbar(aes(ymax = avg_moisture + se_moisture, ymin = avg_moisture - se_moisture), position = dodge, width = 0) +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Fuel moisture (%)") +
  theme(legend.position = "top")
)
ggsave("figures/fuel_moisture_bar.png", height = 4, width = 5, dpi = 300)

```

```{r fuel moisture boxplot}

# ggplot(fuel_mass, aes(water_trt, fuel_moisture)) +
#   # geom_violin(draw_quantiles = c()) +
#   geom_boxplot(aes(color = veg_trt), notch = F, outlier.color = "white") +
#   geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .2)) +
#   # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
#   invasion_fill +
#   invasion_color +
#   theme_bw() +
#   def_theme +
#   xlab("") +
#   ylab("Fuel moisture (%)") +
#   theme(legend.position = "top")

ggplot(filter(agg_plot_data, probe_ht=="0cm"), aes(water_trt, avg_fuel_moisture)) +
  geom_boxplot(aes(color = veg_trt), notch = F, outlier.color = "white") +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .2)) +
  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
  invasion_fill +
  invasion_color +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Fuel moisture (%)") +
  theme(legend.position = "top")

knitr::kable(filter(agg_plot_data, probe_ht=="0cm") %>% 
    group_by(water_trt, veg_trt) %>% 
    summarise(avg_moisture = mean(avg_fuel_moisture),
              nobs = length(avg_fuel_moisture),
              se_moisture = sd(avg_fuel_moisture)/sqrt(nobs)), 
    digits = 2, caption = "Fuel moisture")

```

## Char height

```{r char height fig}
# tree_dmg %>% 
#          group_by(veg_trt, water_trt) %>% 
#          summarise(avg_ht = mean(char_ht_cm),
#                    nobs = length(char_ht_cm),
#                    se = sd(ht2_cm, na.rm = T)/sqrt(nobs))
ggplot(tree_dmg, 
       aes(water_trt, char_ht_cm)) +
  # geom_col(position = dodge) + 
  # geom_errorbar(aes(ymax = avg_ht+se, ymin = avg_ht-se), position = dodge, width = 0) + 
  geom_boxplot(aes(color = veg_trt), alpha = 0) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8)) +
  invasion_fill +
  invasion_color +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Char height (cm)") +
  theme(legend.position = "top")

knitr::kable(tree_dmg %>% 
         group_by(veg_trt, water_trt) %>% 
         summarise(avg_char_ht = mean(char_ht_cm),
                   avg_tree_ht = mean(ht2_cm),
                   nobs = length(char_ht_cm)),
         digits = 2, caption = "Average char & tree heights")

```


## Tree height

```{r tree height fig}
# tree_dmg %>%
         # group_by(veg_trt, water_trt) %>% 
         # summarise(avg_ht = mean(ht2_cm),
         #           nobs = length(ht2_cm),
         #           se = sd(ht2_cm, na.rm = T)/sqrt(nobs)) %>% 
ggplot(tree_dmg, 
       aes(water_trt, ht2_cm)) +
  # geom_col(position = dodge) + 
  # geom_errorbar(aes(ymax = avg_ht+se, ymin = avg_ht-se), position = dodge, width = 0) + 
  geom_boxplot(aes(color = veg_trt)) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .75), alpha = 1) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
  invasion_fill +
  invasion_color +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Tree height (cm)") +
  theme(legend.position = "top")

knitr::kable(tree_dmg %>% 
         group_by(veg_trt, water_trt) %>% 
         summarise(avg_ht = mean(ht2_cm),
                   nobs = length(ht2_cm),
                   se = sd(ht2_cm, na.rm = T)/sqrt(nobs)),
         digits = 2, caption = "Average tree height")

```

```{r tree height lm}

treeht_lm1 <- lm(ht2_cm ~ water_trt*veg_trt, data = tree_dmg)

summary(treeht_lm1)
car::Anova(treeht_lm1, type = "II")

sjPlot::plot_model(treeht_lm1, type = "diag")
sjPlot::plot_model(treeht_lm1, type = "int" , show.data = FALSE)

sjPlot::plot_model(
  treeht_lm1, 
  show.values = T, show.data = T,
  vline.color = "gray90",
  value.size = 5, dot.size = 3.5, line.size = 1, value.offset = .2,
  colors = "viridis",
  title = "Tree height (cm)", 
  # axis.labels = c("invaded","drought*invaded","drought")
  ) +
  theme_bw()+
  theme(axis.text = element_text(size = 12),
        panel.grid = element_blank())

```

The model below with plot as a random effect gives very similar estimates for the treatment effects as the model with no random effect above.

```{r tree height lmer}

treeht_lmer1 <- lmerTest::lmer(ht2_cm ~ water_trt*veg_trt + (1|Plot), data = tree_dmg)
summary(treeht_lmer1)

car::Anova(treeht_lmer1, type = "2")

sjPlot::plot_model(treeht_lmer1, type = "est", show.values = T, vline.color = "gray70") +
  theme_bw()

sjPlot::plot_model(treeht_lmer1, type = "int", show.data = T)

```


## Flame height figures

```{r flame height barchart, eval=F}

ggplot(flame_hts_trt, aes(water_trt, avg_flameht_cm, fill = veg_trt)) +
  # geom_boxplot(aes(fill = veg_trt), notch = F, alpha = .5) +
  # geom_violin(aes(fill = veg_trt), alpha = .5) +
  # geom_point(aes(color = veg_trt), position = "jitter") +
  invasion_color + 
  invasion_fill +
  geom_col(position = dodge) +
  geom_errorbar(aes(ymax = avg_flameht_cm + se_flameht, ymin = avg_flameht_cm - se_flameht), position = dodge, width = 0) +
  # geom_point(aes(color = veg_trt)) +
  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2)) +
  # facet_grid(probe_ht~.) +
  theme_bw() +
  def_theme +
  theme(legend.position = "top") +
  xlab("") +
  ylab("Flame height (cm)")
# ggsave("figures/flame_hts_bar.png", height = 4, width = 5, dpi = 300)

```

```{r flame height boxplot}

ggplot(agg_fuel_flame_data, aes(water_trt, avg_flameht_cm)) +
  geom_boxplot(aes(color = veg_trt), outlier.colour = "white", ) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .4)) +
  invasion_color + 
  invasion_fill +
  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = .8) +
  # facet_grid(probe_ht~.) +
  theme_bw() +
  def_theme +
  theme(legend.position = "top") +
  xlab("") +
  ylab("Flame height (cm)")

```

```{r flame heights table}

knitr::kable(flame_heights_trt, digits = 2)
 
```

Two observations of flame height at three locations in each plot, with ten plots in each treatment group.

# Temperature Data

Two temperature profile examples with a horizontal line indicating 100ºC on each graph.

```{r temperature profile example}

ggplot(filter(temperatures, Plot==8, Box_ID==9),
       aes(time, probe_tempC)) +
  geom_path(aes(color = probe_ht)) +
  geom_hline(yintercept = 100) +
  theme_classic() +
  ggtitle("Cogongrass-invaded with drought treatment")
  NULL
  
ggplot(filter(temperatures, Plot==16, Box_ID==9),
       aes(time, probe_tempC)) +
  geom_path(aes(color = probe_ht)) +
  geom_hline(yintercept = 100) +
  theme_classic() +
  ggtitle("Reference vegetation with drought treatment")
  NULL

```


## Time >100 ºC figures

```{r time above 100 barchart, eval=FALSE}

ggplot(temps_abv_100_data %>% 
         group_by(probe_ht, veg_trt, water_trt) %>% 
         summarise(avg_s_abv100 = mean(seconds_abv_100),
                   nobs = length(seconds_abv_100),
                   se_s_abv100 = sd(seconds_abv_100/sqrt(nobs))), 
       aes(water_trt, avg_s_abv100, fill = veg_trt)) +
  geom_col(position = dodge) +
  geom_errorbar(aes(ymax = avg_s_abv100 + se_s_abv100, ymin = avg_s_abv100 - se_s_abv100), position = dodge, width = 0) +
  invasion_fill +
  facet_grid(.~probe_ht) +
  theme_bw() + 
  def_theme +
  ylab("Seconds >100 ºC") +
  xlab("") +
  theme(legend.position = "top")
# ggsave("figures/seconds_abv100_bar.png", height = 4, width = 9, dpi = 300)

```


```{r temps above 100 figs}

# ggplot(filter(temps_abv_100_data, seconds_abv_100<1200), aes((drywt*16)/1000, seconds_abv_100)) +
#   geom_point(aes(color = veg_trt, shape = water_trt), alpha = .5) +
#   # stat_summary(aes(color = veg_trt), fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
#   # geom_boxplot(aes(color = veg_trt)) +
#   # scale_y_continuous(breaks = c(0, 100, 150, 200), labels = c("0","100","150","200")) +
#   geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm", show.legend = F) +
#   facet_grid(probe_ht~.) +
#   invasion_color +
#   invasion_fill +
#   theme_bw() +
#   def_theme +
#   ylab("Seconds >100 ºC") +
#   xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
#   theme(legend.position = "top",
#         legend.direction = "vertical") +
#   NULL
# ggsave("figures/seconds_abv100.png", height = 4, width = 9, dpi = 300)

## Temperature metrics aggregated to n=10 plots for each treatment combo
fig_labels <- data.frame(
  probe_ht = c("50cm","25cm","0cm"), 
  panel_id = c("a","b","c"),
  avg_dry_wt=c(-.1,-.1,-.1), 
  avg_s_abv100=c(450,450,450))

time_abv_100_fig <- ggplot(agg_plot_data, aes((avg_dry_wt*16)/1000, avg_s_abv100)) +
  geom_point(aes(color = veg_trt, shape = water_trt)) +
  scale_shape_discrete(solid = FALSE) +
  invasion_color +
  invasion_fill +
  geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm", show.legend = FALSE) +
  geom_text(data = fig_labels, aes(label = panel_id), fontface = "bold") +
  facet_grid(probe_ht~.) +
  theme_bw() +
  # theme_classic() +
  def_theme +
  theme(legend.direction = "vertical",
        legend.background = element_blank()) +
  xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
  ylab("Temperature residence time (s >100ºC)")

# time_abv_100_fig

## heat flux ####
# ggplot(agg_plot_data, aes((avg_dry_wt*16)/1000, avg_heat_flux100)) +
#   geom_point(aes(color = veg_trt, shape = water_trt), alpha = .5) +
#   invasion_color +
#   invasion_fill +
#   geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm") +
#   facet_grid(probe_ht~.) +
#   theme_bw() +
#   def_theme +
#   theme(legend.position = c(.2,.9), 
#         legend.direction = "vertical",
#         legend.background = element_blank()) +
#   xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
#   ylab("Heat flux >100 ºC (s * temperature)") +
#   NULL


# ggplot(temps_abv_100_data, aes(avg_drywt, seconds_abv_100)) +
#   geom_point(aes(color = veg_trt)) +
#   geom_smooth(aes(color = veg_trt), se = F) +
#   facet_grid(.~probe_ht, scales = "free_y") +
#   theme_bw() +
#   def_theme

```

```{r temps abv 100 table}

knitr::kable(temps_abv_100_data %>% 
         group_by(probe_ht, veg_trt, water_trt) %>% 
         summarise(avg_s_abv100 = mean(seconds_abv_100),
                   nobs = length(seconds_abv_100),
                   se_s_abv100 = sd(seconds_abv_100/sqrt(nobs))),
         digits = 2, caption = "Seconds above 100 ºC by treatment and probe height")

```

## Max temp figures

```{r max temp barchart, eval=FALSE}

ggplot(max_temperatures %>% 
         group_by(probe_ht, water_trt, veg_trt) %>% 
         summarise(avg_max_temp = mean(max_temp),
                   nobs = length(max_temp),
                   se_max_temp = sd(max_temp)/sqrt(nobs)), 
       aes(water_trt, avg_max_temp, fill = veg_trt)) +
  
  # stat_summary(aes(color = veg_trt), 
               # fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2)) +
  geom_col(position = dodge) +
  geom_errorbar(aes(ymax = avg_max_temp + se_max_temp, ymin = avg_max_temp - se_max_temp), position = dodge, width = 0) +
  invasion_fill +
  facet_grid(.~probe_ht) +
  theme_bw() + 
  def_theme +
  ylab("Maximum temperature (ºC)") +
  xlab("") +
  theme(legend.position = "top",
        legend.direction = "vertical") +
  NULL
# ggsave("figures/max_temp_bars.png", height = 4, width = 9, dpi = 300)

```

```{r max temp mean_se, eval=FALSE}

ggplot(max_temperatures, aes(water_trt, max_temp)) +
  stat_summary(aes(color = veg_trt), fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
  invasion_color +
  facet_grid(.~probe_ht) +
  theme_bw() + 
  def_theme +
  ylab("Maximum temperature (ºC)") +
  xlab("") +
  theme(legend.position = "top")
# ggsave("figures/max_temp.png", height = 4, width = 9, dpi = 300)

```

The first figure is the temperature and fuel load aggregated (averaged) for each plot, so n=10 for each treatment combination. 
```{r max temp regression figs}

fig_labels <- data.frame(
  probe_ht = c("50cm","25cm","0cm"), 
  panel_id = c("d","e","f"),
  avg_dry_wt=c(-.1,-.1,-.1), 
  avg_max_temp=c(925,925,925))
## Aggregated to averages for n=10 plots for each treatment combination 
max_temp_fig <- ggplot(agg_plot_data, aes((avg_dry_wt*16)/1000, avg_max_temp)) +
  geom_point(aes(color = veg_trt, shape = water_trt), alpha = 1) +
  scale_shape_discrete(solid = FALSE) +
  invasion_color +
  invasion_fill +
  geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm", show.legend = FALSE) +
  geom_text(data = fig_labels, aes(label = panel_id), fontface = "bold") +
  facet_grid(probe_ht~.) +
  theme_bw() +
  # theme_classic() +
  def_theme +
  theme(legend.position = "none", 
        legend.direction = "vertical",
        legend.background = element_blank()) +
  xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
  ylab("Maximum temperature (ºC)")
# max_temp_fig
# ggsave("figures/max_temp_scatter.png", height = 4, width = 9, dpi = 300)

maxt_mod <- (lm(avg_max_temp ~ avg_drywt + veg_trt, data = agg_plot_data))
summary(maxt_mod)

sjPlot::plot_model(maxt_mod, type = "pred", terms = c("avg_drywt","veg_trt"), show.data = T)

## Not aggregated, so n=36 (max) for each treatment combination
# ggplot(max_temperatures, aes((drywt*16)/1000, max_temp_avg5)) +
#   geom_point(aes(color = veg_trt), alpha = .5) +
#   invasion_color +
#   invasion_fill +
#   geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm") +
#   facet_grid(probe_ht~water_trt) +
#   theme_bw() +
#   def_theme +
#   theme(legend.position = c(.396,.275), 
#         legend.direction = "vertical",
#         legend.background = element_blank()) +
#   xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
#   ylab("Maximum temperature (ºC)")

```

```{r max temp summary tables}

knitr::kable(max_temperatures %>% 
         group_by(probe_ht, water_trt, veg_trt) %>% 
         summarise(avg_max_temp = mean(max_temp),
                   nobs = length(max_temp),
                   se_max_temp = sd(max_temp)/sqrt(nobs)), 
         digits = 2, caption = "Maximum temperature by treatment and probe height")

knitr::kable(max_temperatures %>% 
  group_by(probe_ht, veg_trt) %>% 
  summarise(avg_max_temp = mean(max_temp),
            nobs = length(max_temp),
            se_max_temp = sd(max_temp)/sqrt(nobs),
            maximum_temp = max(max_temp)),
  digits = 2, caption = "Maximum temperature by invasion status and probe height")

```

## Temperatures manuscript figure

```{r fig1 }

cowplot::plot_grid(
  time_abv_100_fig +
    theme(strip.text = element_text(color = "white"),
          legend.margin = margin(t = -3, r = 0, b = -3, l = 0),
          legend.position = c(.7,.9),),
  max_temp_fig +
    theme(strip.background = element_blank()))

# ggsave("figures/fig1_7x7_600dpi.tiff", height = 7, width = 7, dpi = 600)
# Recommended max height: 54 picas / 9” / 22.5 cm. 
# Use one of the following widths: 
# 1 column wide (20.5 picas / 3.42” / 8.7 cm) 
# 1.5 columns wide (27 picas / 4.5” / 11.4 cm) 
# 2 columns wide (42.125 picas / 7” / 17.8 cm)

```


## Max temp regressions
```{r max temp 50cm regressions}

max_temps50_data <- max_temperatures %>% 
  filter(probe_ht=="50cm") %>% 
  mutate(veg_trt = as.factor(veg_trt),
         water_trt = as.factor(water_trt))
summary(max_temps50_data)

max_temp50_lme1 <- lmerTest::lmer(max_temp ~ drywt*veg_trt + drywt*water_trt + fuel_moisture + veg_trt*water_trt + (1|Plot), data = max_temps50_data, na.action = "na.fail")
summary(max_temp50_lme1)
MuMIn::r.squaredGLMM(max_temp50_lme1)
# dmod <- MuMIn::dredge(max_temp50_lme1)
# car::Anova(max_temp50_lme1, type = "II", test = "F")

max_temp50_lme2 <- lmerTest::lmer(max_temp ~ drywt + veg_trt + fuel_moisture + water_trt + (1|Plot), data = max_temps50_data)
summary(max_temp50_lme2)
anova(max_temp50_lme1,max_temp50_lme2)
car::Anova(max_temp50_lme2)

max_temp50_lme3 <- lmerTest::lmer(max_temp ~ drywt + veg_trt + water_trt + avg_fuel_ht_cm + (1|Plot), data = max_temps50_data)
summary(max_temp50_lme3)
car::Anova(max_temp50_lme3)

max_temp50_lme4a <- lmerTest::lmer(max_temp ~ drywt*veg_trt + avg_fuel_ht_cm + (1|Plot), data = max_temps50_data)
summary(max_temp50_lme4a)
car::Anova(max_temp50_lme4a)

max_temp50_lme4 <- lme4::lmer(max_temp ~ drywt + veg_trt + (1|Plot), data = max_temps50_data)
summary(max_temp50_lme4)
car::Anova(max_temp50_lme4, type = "II")

sjPlot::plot_model(
  max_temp50_lme4, type = "est", 
  show.values = T, vline.color = "gray90", colors = "viridis") +
  theme_bw()



# sjPlot::plot_model(max_temp50_lme4, type = "pred", terms = c("avg_drywt","veg_trt"), show.data = T)

# min(max_temps50_data$avg_drywt)*1.14 + 190
# median(max_temps50_data$avg_drywt)*1.14 + 190
# max(max_temps50_data$avg_drywt)*1.14 + 190
# min(max_temps50_data$avg_drywt)*1.14 + 190 + 328.26
# median(max_temps50_data$avg_drywt)*1.14 + 190 + 328.26
# max(max_temps50_data$avg_drywt)*1.14 + 190 + 328.26

sjPlot::plot_model(
  max_temp50_lme4, type = "std2", show.values = T, 
  vline.color = "black", colors = "viridis",
  title = "Standardized coefficients for max temp (divided by 2*sd)") +
  theme_bw()
MuMIn::r.squaredGLMM(max_temp50_lme4)

lm1 <- lm(max_temp ~ drywt + veg_trt, data = max_temps50_data)
summary(lm1)

sjPlot::plot_model(lm1, type = "pred", terms = c("drywt","veg_trt"), show.data = TRUE) +
  invasion_color +
  invasion_fill +
  theme_classic()

# max_temps50_data_plot <- max_temps50_data %>% 
#   group_by(Plot, water_trt, veg_trt) %>% 
#   summarise(avg_max_temp = mean(max_temp),
#             avg_drywt = mean(drywt),
#             avg_moisture = mean(fuel_moisture),
#             avg_fuel_ht_cm = mean(avg_fuel_ht_cm)) %>% 
#   ungroup(.)

# plot(max_temps50_lm)


max_temps50_lm <- lm(avg_max_temp ~ avg_drywt + veg_trt, data = filter(agg_plot_data, probe_ht=="50cm"))
summary(max_temps50_lm)

ggplot(filter(agg_plot_data, probe_ht=="50cm"), aes(.001*avg_drywt*16, avg_max_temp)) +
  geom_point(aes(color = veg_trt, shape = water_trt)) +
  geom_smooth(aes(fill = veg_trt, color = veg_trt), method = "lm", show.legend = FALSE) +
  scale_shape_discrete(solid = FALSE) +
  invasion_color +
  invasion_fill +
  theme_classic() +
  xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
  ylab("Maximum temperature at 50 cm (ºC)") +
  theme(legend.title = element_blank()) +
  NULL
  
```

```{r max temp 25cm regressions}

max_temps25_data <- max_temperatures %>% 
  filter(probe_ht=="25cm")

max_temp25_lme4 <- lme4::lmer(max_temp ~ drywt*veg_trt + (1|Plot), data = max_temps25_data)
summary(max_temp25_lme4)
car::Anova(max_temp25_lme4, type = "II")
MuMIn::r.squaredGLMM(max_temp25_lme4)

sjPlot::plot_model(
  max_temp25_lme4, type = "est", 
  show.values = T, vline.color = "gray90", colors = "viridis") +
  theme_bw()


```

```{r max temp 0cm regressions}

max_temps0_data <- max_temperatures %>% 
  filter(probe_ht=="0cm")

max_temp0_lme4 <- lme4::lmer(max_temp ~ drywt + veg_trt + (1|Plot), data = max_temps0_data)
summary(max_temp0_lme4)
car::Anova(max_temp0_lme4, type = "II")
MuMIn::r.squaredGLMM(max_temp0_lme4)

sjPlot::plot_model(
  max_temp0_lme4, type = "est", 
  show.values = T, vline.color = "gray90", colors = "viridis") +
  theme_bw()

```

## Time >100 ºC regressions: 50cm probe height
```{r time abv 100 regressions}

temps_abv100_50data <- temps_abv_100_data %>% 
  filter(probe_ht=="50cm")

abv100_lme1 <- lmerTest::lmer(seconds_abv_100 ~ drywt*veg_trt + fuel_moisture + avg_litter_depth_cm*veg_trt + veg_trt*water_trt + (1|Plot), data = temps_abv100_50data)
summary(abv100_lme1)

abv100_lm <- lm(seconds_abv_100 ~ drywt*veg_trt + avg_litter_depth_cm*veg_trt, data = temps_abv100_50data)
summary(abv100_lm)


abv100_lme2 <- lmerTest::lmer(seconds_abv_100 ~ drywt*veg_trt + fuel_moisture + water_trt + (1|Plot), data = temps_abv100_50data)
summary(abv100_lme2)

abv100_lme3 <- lmerTest::lmer(seconds_abv_100 ~ drywt*veg_trt + fuel_moisture + (1|Plot), data = temps_abv100_50data)
summary(abv100_lme3)

abv100_lme4 <- lmerTest::lmer(seconds_abv_100 ~ drywt*veg_trt + (1|Plot), data = temps_abv100_50data)
summary(abv100_lme4)

plot(abv100_lme4)

sjPlot::plot_model(abv100_lme4, type = "est", show.values = T, vline.color = "gray90", colors = "viridis") +
  theme_bw()
MuMIn::r.squaredGLMM(abv100_lme4)

```

# Tree Damage & Survival


## Percent dead limbs

```{r pct dead limbs boxplot}

ggplot(tree_dmg, aes(water_trt, pct_dead_limbs)) +
  # stat_summary(aes(fill = veg_trt), fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .5), size = 1, shape = 21) +
  geom_boxplot(aes(fill = veg_trt), alpha = 0, show.legend = FALSE) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .8, jitter.width = .2), show.legend = TRUE) +
  invasion_color +
  invasion_fill +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Dead limbs (%)") +
  theme(legend.position = "top")
# ggsave("figures/pct_dead_limbs.png", height = 4, width = 5, dpi = 300)

```

```{r pct dead limbs barchart, eval = FALSE}

ggplot(tree_dmg %>% 
         group_by(veg_trt, water_trt) %>% 
         summarise(avg_pct_dead_limbs = mean(pct_dead_limbs),
                   nobs = length(pct_dead_limbs),
                   se_dead_limbs = sd(pct_dead_limbs, na.rm = T)/sqrt(nobs)), 
       aes(water_trt, avg_pct_dead_limbs, fill = veg_trt)) +
  geom_col(position = dodge) + 
  geom_errorbar(aes(ymax = avg_pct_dead_limbs+se_dead_limbs, ymin = avg_pct_dead_limbs-se_dead_limbs), position = dodge, width = 0) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2)) +
  # invasion_color +
  invasion_fill +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("Dead limbs (%)") +
  theme(legend.position = "top")
# ggsave("figures/pct_dead_limbs_bar.png", height = 4, width = 5, dpi = 300)

```

```{r pct dead limbs data table summary}

knitr::kable(tree_dmg %>% 
         group_by(veg_trt, water_trt) %>% 
         summarise(avg_pct_dead_limbs = mean(pct_dead_limbs),
                   nobs = length(pct_dead_limbs),
                   se_dead_limbs = sd(pct_dead_limbs, na.rm = T)/sqrt(nobs))
         )

```

## Percent canopy brown

The same pattern as percent dead limbs.

```{r pct brown boxplot}

ggplot(tree_dmg, aes(water_trt, pct_brown)) +
  # stat_summary(aes(fill = veg_trt), fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .5), size = 1, shape = 21) +
  geom_boxplot(aes(fill = veg_trt), alpha = 0, show.legend = FALSE) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .75, jitter.width = .3), show.legend = TRUE, alpha = 1) +
  invasion_color +
  invasion_fill +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("% Brown") +
  theme(legend.position = "top")
# ggsave("figures/pct_brown.png", height = 4, width = 5, dpi = 300)

```

```{r pct brown data table summary}

knitr::kable(tree_dmg %>% 
         group_by(veg_trt, water_trt) %>% 
         summarise(avg_pct_brown = mean(pct_brown),
                   nobs = length(pct_dead_limbs),
                   se_pct_brown = sd(pct_brown, na.rm = T)/sqrt(nobs)))

```

## Post-fire tree survival

Box-plots showing the proportion of trees in each plot that survived the fires.

```{r llp survival}

tree_surv_plot <- tree_dmg %>% 
  group_by(Plot, water_trt, veg_trt) %>% 
  summarise(survival = sum(survival)/length(survival))

ggplot(tree_surv_plot, aes(water_trt, survival*100)) +
  geom_boxplot(aes(fill = veg_trt), alpha = 0, show.legend = FALSE) +
  geom_point(aes(color = veg_trt), position = position_jitterdodge(dodge.width = .75, jitter.width = .3)) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .5), size = 1) +
  invasion_fill +
  invasion_color +
  theme_bw() +
  def_theme +
  xlab("") +
  ylab("% survival") +
  theme(legend.position = "top")

```

```{r tree survival data table summary}

tree_dmg %>% 
  group_by(water_trt, veg_trt) %>% 
  summarise(plots = n_distinct(Plot),
            trees = length(tree_id),
            pct_survival = 100*sum(survival)/length(survival),
            sd_survival = 100*(sd(survival, na.rm = TRUE))) %>% 
  knitr::kable(.)

```

```{r plot-level tree mortality models, eval=FALSE}

# library(lme4)
# library(arm)# standardize() function centers & divide 2*sd
# library(piecewiseSEM)

## df of 19 trees in the ambientXinvasion treatment
# tree_dmg2 <- tree_dmg %>% 
  # filter(veg_trt != "reference", water_trt != "drought")

# tree_dmg2 <- anti_join(tree_dmg,tree_dmg2)

surv_glm1 <- glm(cbind(trees_alive,trees_dead) ~ avg_tree_ht_cm*veg_trt + avg_max_temp + avg_flameht_cm + veg_trt*water_trt, family = "binomial", data = filter(agg_plot_data, probe_ht=="50cm"))
summary(surv_glm1)
sjPlot::plot_model(surv_glm1, type = "pred", terms = c("avg_max_temp","veg_trt"))
# car::Anova(surv_glm1, type = "III", test = "LR")
## max temperature not significant
## s above 100 not significant
## heat flux not significant

surv_glm2 <- glm(cbind(trees_alive,trees_dead) ~ avg_tree_ht_cm + avg_heat_flux100 + veg_trt*water_trt, family = "binomial", data = filter(agg_plot_data, probe_ht=="25cm"))
summary(surv_glm2)
sjPlot::plot_model(surv_glm2, type = "pred", terms = c("avg_heat_flux100","veg_trt"))
## max temperature not significant
## s above 100 not significant
## heat flux not significant

surv_glm3 <- glm(cbind(trees_alive,trees_dead) ~ avg_tree_ht_cm + avg_heat_flux100 + veg_trt*water_trt, family = "binomial", data = filter(agg_plot_data, probe_ht=="0cm"))
summary(surv_glm3)
sjPlot::plot_model(surv_glm3, type = "pred", terms = c("avg_tree_ht_cm","veg_trt"))
sjPlot::plot_model(surv_glm3, type = "pred", terms = c("avg_heat_flux100","veg_trt"))
## max temperature not significant
## s above 100 not significant
## heat flux not significant


tree_dmg <- tree_dmg %>% 
  mutate(char_ht_scaled = scale(char_ht_cm, scale = 2*sd(char_ht_cm)),
         ht2_scaled = scale(ht2_cm, scale = 2*sd(ht2_cm)))

surv_glmer1 <- lme4::glmer(cbind(alive,dead) ~ char_ht_scaled + ht2_scaled + veg_trt + water_trt + (1|Plot), family = "binomial", data = tree_dmg)
# sjPlot::plot_model(surv_glmer1, type="diag")
sjPlot::plot_model(surv_glmer1)
summary(surv_glmer1)
car::Anova(surv_glmer1, type = "II")
```

```{r pairs plot, eval=FALSE}

tree_agg_plot_data %>% 
  ungroup(.) %>% 
  filter(probe_ht=="0cm") %>% 
  select(survival, avg_max_temp:avg_flameht_cm, char_ht_cm, ht2_cm, veg_trt, water_trt) %>% 
  GGally::ggpairs(., aes(color = veg_trt))

```

```{r tree-level mortality modeling}

## Which probe height temperature measurement best explains tree survival? ####

# write_csv(tree_agg_plot_data,"data/tree_mortality_model_data.csv")
tree_agg_plot_data <- read_csv("data/tree_mortality_model_data.csv")

## Rescale variables
### Continuous: center and divide by 2 standard deviations
model_data <- tree_agg_plot_data %>% 
  # mutate(tree_id = paste(Plot, ))
  group_by(probe_ht) %>%
  # ungroup(.) %>%
  # filter(probe_ht=="50cm") %>% 
  # select(survival, avg_max_temp:avg_flameht_cm, char_ht_cm, ht2_cm, veg_trt, water_trt, Plot, tree_id, Ind.) %>% 
  mutate(
    max_temp_scl = (avg_max_temp-mean(avg_max_temp))/(2*sd(avg_max_temp)),
    maxt_scale = scale(avg_max_temp),
    
    s_abv100_scl = (avg_s_abv_100-mean(avg_s_abv_100))/(2*sd(avg_s_abv_100)),
    sabv100_scale = scale(avg_s_abv_100),
         
    heat_flux_scl = (avg_heat_flux100-mean(avg_heat_flux100))/(2*sd(avg_heat_flux100)),
    hflux_scale = scale(avg_heat_flux100),
    
    dry_wt_scl = (avg_drywt-mean(avg_drywt))/(2*sd(avg_drywt)),
    wt_scale = scale(avg_drywt),
    
    fuel_ht_scl = (avg_fuel_ht_cm-mean(avg_fuel_ht_cm))/(2*sd(avg_fuel_ht_cm)),
    fuelht_scale = scale(avg_fuel_ht_cm),
    
    tree_ht_scl =  (ht2_cm-mean(ht2_cm))/(2*sd(ht2_cm)),
    treeht_scale = scale(ht2_cm),
    
    flame_ht_scl =  (avg_flameht_cm-mean(avg_flameht_cm))/(2*sd(avg_flameht_cm)),
    flameht_scale = scale(avg_flameht_cm),
    
    char_ht_scl = (char_ht_cm-mean(char_ht_cm))/(2*sd(char_ht_cm)),
    charht_scale = scale(char_ht_cm),
    
    veg_trt = recode_factor(veg_trt, "reference" = 0, "invaded" = 1),
    water_trt = recode_factor(water_trt, "ambient" = 0 , "drought" = 1),
    
    Treatment = factor(Treatment, levels = c("AN","AC","DN","DC"))
    
         ) %>% 
  ungroup(.) %>% 
  mutate(Plot = as.factor(Plot),
         tree_id = paste(Plot, Ind., sep = ""))
# summary(model_data)

dat0cm <- droplevels(filter(model_data, probe_ht=="0cm"))
dat25cm <- droplevels(filter(model_data, probe_ht=="25cm"))
dat50cm <- droplevels(filter(model_data, probe_ht=="50cm"))

# summary(select(dat0cm, tree_ht_scl, heat_flux_scl, flame_ht_scl))
# summary(select(dat25cm, tree_ht_scl, heat_flux_scl, flame_ht_scl))
# summary(select(dat50cm, tree_ht_scl, heat_flux_scl, flame_ht_scl))

library(lme4)

m0cm <- glmer(
  survival ~ tree_ht_scl + 
    flame_ht_scl +
    # char_ht_scl + #AIC: 70, "boundary (singular) fit"
    max_temp_scl + #AIC: 66.9
    # s_abv100_scl + #AIC: 66.9, "boundary (singular) fit"
    # heat_flux_scl + #AIC: 67
    (1|Plot),
  family = "binomial", data = dat0cm)
# summary(m0cm)
# sjPlot::plot_model(m0cm, type = "diag")
# MuMIn::r.squaredGLMM(m0cm)

m25cm <- glmer(
  survival ~ tree_ht_scl + 
    flame_ht_scl +
    # char_ht_scl + #AIC 63.7, "boundary (singular) fit"
    max_temp_scl + #AIC: 64.8
    # s_abv100_scl + #AIC: 66.9
    # heat_flux_scl + #AIC: 66.9
    (1|Plot), 
  family = "binomial", data = dat25cm)
# summary(m25cm)
# sjPlot::plot_model(m25cm, type = "diag")
# MuMIn::r.squaredGLMM(m25cm)

m25cm_trt <- glmer(survival ~ veg_trt*water_trt + (1|Plot), data = dat25cm,
                   family = "binomial")
summary(m25cm_trt)


m50cm <- glmer(
  survival ~ tree_ht_scl + 
    flame_ht_scl +
    # char_ht_scl + #AIC: 63.9, "boundary (singular) fit"
    max_temp_scl + #AIC: 64.8, "boundary (singular) fit"
    # s_abv100_scl + #AIC: 67.0
    # heat_flux_scl + #AIC: 66.1
    (1|Plot), 
  family = "binomial", data = dat50cm)
# summary(m50cm)
# sjPlot::plot_model(m50cm, type = "diag")
# MuMIn::r.squaredGLMM(m50cm)
# 
# AIC(m0cm,m25cm,m50cm)
## Based on comparisons of the above models using AIC 25 or 50 cm is best
## 50 cm produced singular fit, so opting for 25 cm

m50cm_stan <- stan_glmer(survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1|Plot), family = binomial(link="logit"), data = dat50cm)
summary(m50cm_stan)

m25cm_unscaled <- glmer(
  survival ~ ht2_cm + avg_flameht_cm + avg_max_temp + (1|Plot),
  family = "binomial", data = dat25cm,
  # control = glmerControl(
    # optimizer="bobyqa", optCtrl = list(maxfun=2e5)
    # )
            )
# summary(m25cm_unscaled)# Convergence failure warning: "rescale variables"
## Use unscaled model for some visualizations
# summary(m25cm)

```



```{r testing models, eval=FALSE}

surv_glmer0cm <- lme4::glmer(
  survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1|Plot),
  family = "binomial", 
  # control = glmerControl(optCtrl = list(maxfun=2e5)),
  data = dat0cm)
summary(surv_glmer0cm)
car::Anova(surv_glmer0cm, type = "II")

m1 <- glmer(survival ~ tree_ht_scl + flame_ht_scl + heat_flux_scl +(1|Plot), family = "binomial", data = dat0cm,
            # control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun=2e5))
            )
summary(m1)

m2 <- glmer(
  survival ~ ht2_cm + avg_flameht_cm + avg_max_temp + (1|Plot),
  family = "binomial", data = dat0cm,
  # control = glmerControl(
    # optimizer="bobyqa", optCtrl = list(maxfun=2e5)
    # )
            )
summary(m2)

# m3 <- lmerTest::lmer(
#   pct_brown ~ ht2_cm + char_ht_cm + avg_max_temp + (1|Plot), data = dat0cm)
# summary(m3)


surv_glmer25cm <- lme4::glmer(
  survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1|Plot), 
  family = "binomial", data = dat25cm)
summary(surv_glmer25cm)
# arm::standardize(surv_glmer25cm)
m4 <- glmer(
  survival ~ ht2_cm + avg_max_temp + avg_flameht_cm + (1|Plot),
  family = "binomial", data = dat25cm,
  # control = glmerControl(
    # optimizer="bobyqa", optCtrl = list(maxfun=2e5)
    # )
            )
summary(m4)

surv_glmer50cm <- lme4::glmer(
  survival ~ tree_ht_scl + max_temp_scl + flame_ht_scl + (1|Plot),
  # control = glmerControl(optimizer = "bobyqa"),
  family = "binomial", data = dat50cm 
  )
summary(surv_glmer50cm)

AIC(surv_glmer0cm, surv_glmer25cm, surv_glmer50cm)
# surv_glmer50cm <- lme4::glmer(survival ~ ht2_cm*veg_trt + avg_heat_flux100 + avg_flameht_cm + (1|Plot), family = "binomial", data = model_data)
# summary(surv_glmer50cm)
# surv_glmer50cm.std <- lme4::glmer(survival ~ tree_ht_scl + heat_flux_scl + flame_ht_scl + (1|Plot), family = "binomial", data = model_data)

sjPlot::plot_model(surv_glmer25cm, type = "pred", terms = c("max_temp_scl", "flame_ht_scl", "tree_ht_scl"), pred.type = "re" )
sjPlot::plot_model(m4, type = "pred", terms = c("avg_max_temp", "avg_flameht_cm","ht2_cm"), pred.type = "fe", show.data = TRUE)
sjPlot::plot_model(m4, type = "est", show.values = T)

```

Build separate models for the various relationships of interest using subsets of the data as appropriate for response variables aggregated to the plot such as temperature and flame height

```{r piece models together}

library(nlme)
library(sjPlot)

library(rstanarm)

options(mc.cores = parallel::detectCores())

surv1 <- stan_glmer(survival ~ veg_trt*water_trt + (1|Plot), data = dat25cm,
                   family = binomial(link = "logit"))
print(surv1, digits = 3)


m25cm1 <- glmer(
  survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1|Plot), 
  family = binomial(link = logit), data = dat25cm)
summary(m25cm1)

m1 <- stan_glmer(survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1 | Plot), data = dat25cm, family = binomial(link = "logit"), prior = student_t(4,0,2.5))
print(m1, digits = 3)
plot(m1)


m50cm1 <- glmer(
  survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1|Plot), 
  family = binomial(link = logit), data = dat50cm)
summary(m50cm1)

m2 <- stan_glmer(survival ~ tree_ht_scl + flame_ht_scl + max_temp_scl + (1 | Plot), data = dat50cm, family = binomial(link = "logit"), prior = student_t(4,0,2.5))
print(m2, digits = 3)


## Plot Level Models ####
tree_plots_dat <- filter(agg_plot_data, Plot %in% unique(dat25cm$Plot))

## Max Temp Model

max_temp_fig <- ggplot(tree_plots_dat, aes((avg_drywt*16)/1000, avg_max_temp)) +
  geom_point(aes(color = veg_trt, shape = water_trt), alpha = 1) +
  scale_shape_discrete(solid = FALSE) +
  invasion_color +
  invasion_fill +
  geom_smooth(aes(color = veg_trt, fill = veg_trt), alpha = .2, method = "lm", show.legend = FALSE) +
  # geom_text(data = fig_labels, aes(label = panel_id), fontface = "bold") +
  facet_grid(probe_ht~.) +
  theme_bw() +
  # theme_classic() +
  def_theme +
  theme(legend.position = "none", 
        legend.direction = "vertical",
        legend.background = element_blank()) +
  xlab(expression(paste("Fuel load (kg ", {m}^-2, ")"), sep = "")) +
  ylab("Maximum temperature (ºC)")

maxtemp25cm <- lm(avg_max_temp ~ avg_drywt + veg_trt, data = filter(tree_plots_dat, probe_ht == "25cm"))
summary(maxtemp25cm)

maxtemp50cm <- lm(avg_max_temp ~ avg_drywt + veg_trt, data = filter(tree_plots_dat, probe_ht == "50cm"))
summary(maxtemp50cm)


sabv100_50cm <- lm(avg_s_abv_100 ~ avg_drywt*veg_trt, data = filter(tree_plots_dat, probe_ht == "50cm"))
summary(sabv100_50cm)

hflux_50cm <- lm(avg_heat_flux100 ~ avg_drywt + veg_trt, data = filter(tree_plots_dat, probe_ht == "50cm"))
summary(hflux_50cm)


```


### Build path model leading to tree mortality

Use rescaled variables in the path model to understand the relative effects. The rescaled variables are centered (subtracting the mean) and then divided by 2 standard deviations to put the continuos variables on the same scale as the categorical variables. 


```{r build path model}
## Tree mortality model
# m25cm
# m25cm_unscaled

plot_model(m25cm, show.p = T, show.values = T)

plot_model(m25cm, type = "eff", show.data = T)
plot_model(m25cm_unscaled, type = "eff", show.data = T)

maxtemp25cm <- lme(avg_max_temp ~ avg_drywt, random = ~1|Plot, 
                    data = dat25cm)
summary(maxtemp25cm)
plot_model(maxtemp25cm, type = "diag")
plot_model(maxtemp25cm, type = "eff")
# AIC(maxtemp25cm)

maxt25cm <- lmer(max_temp_scl ~ dry_wt_scl + (1|Plot), 
                    data = dat25cm,
                    # control = lmerControl(
                      # optimizer = "bobyqa",
                      # optCtrl = list(maxfun=2e6)
                    # )
                 )
summary(maxt25cm)
plot_model(maxt25cm, type = "diag")
plot_model(maxt25cm, type = "eff", pred.type = "re", show.data = T)
plot_model(maxt25cm, type = "pred", show.data = T)
plot_model(maxt25cm, show.data = T)
MuMIn::r.squaredGLMM(maxt25cm)


maxt25cm1 <- lmer(avg_max_temp ~ avg_drywt + (1|Plot), 
                    data = dat25cm,
                    # control = lmerControl(
                      # optimizer = "bobyqa",
                      # optCtrl = list(maxfun=2e6)
                    # )
                 )
summary(maxt25cm1)
plot_model(maxt25cm1, type = "diag")
plot_model(maxt25cm1, type = "pred", show.data = T, pred.type = "re")
MuMIn::r.squaredGLMM(maxt25cm1)



flame_ht <- lme(avg_flameht_cm ~ avg_drywt, random = ~ 1|Plot, data = dat25cm)
summary(flame_ht)
plot_model(flame_ht, type = "diag")

flame_ht1 <- lmer(flame_ht_scl ~ dry_wt_scl + (1|Plot), data = dat25cm)
summary(flame_ht1)
plot_model(flame_ht1, type = "pred", show.data = T, pred.type = "re")
MuMIn::r.squaredGLMM(flame_ht1)



fuel_load <- lme(avg_drywt ~ water_trt*veg_trt, random = ~1|Plot, data = dat25cm)
summary(fuel_load)
plot_model(fuel_load)
plot_model(fuel_load, type = "diag")

fuel_load1 <- lmer(dry_wt_scl ~ water_trt*veg_trt + (1|Plot), data = dat25cm)
summary(fuel_load1)
plot_model(fuel_load1)
plot_model(fuel_load1, type = "diag")
plot_model(fuel_load1, type = "int")

tree_ht <- nlme::lme(ht2_cm ~ water_trt*veg_trt , random = ~ 1|Plot, data = dat25cm)
summary(tree_ht)
plot_model(tree_ht)
# tree_ht50cm.std <- standardize(tree_ht50cm)
tree_ht1 <- lmer(tree_ht_scl ~ water_trt*veg_trt + (1|Plot), data = dat25cm)
summary(tree_ht1)
plot_model(tree_ht1)

# devtools::install_github("jslefche/piecewiseSEM@devel", build_vignette=TRUE)
library(piecewiseSEM)
psem.mod <- psem(
  m25cm_unscaled,
  maxtemp25cm,
  flame_ht,
  fuel_load,
  tree_ht
)
summary(psem.mod)


```

Too many issues with the catergorical variables (?) to achieve a fit using SEM methods.


```{r lavaan model, eval = F}
library(lavaan)

myModel <- 'survival ~ ht2_cm + avg_flameht_cm + avg_max_temp
                  avg_max_temp ~ avg_drywt
                  avg_flameht_cm ~ avg_drywt
                  avg_drywt ~ water_trt * veg_trt
                  ht2_cm ~ water_trt * veg_trt

'

fit <- sem(myModel, data=dat25cm)
summary(fit, standardized=TRUE)

  
' # regressions
             

             # latent variable definitions 
               # f1 =~ y1 + y2 + y3 
               # f2 =~ y4 + y5 + y6 
               # f3 =~ y7 + y8 + y9 + y10

             # variances and covariances 
               # y1 ~~ y1 
               # y1 ~~ y2 
               # f1 ~~ f2

             # intercepts 
               # y1 ~ 1 
               # f1 ~ 1
'

```


```{bash, eval=F}
## run manually 
cp  data-summary.html ~/Dropbox\ \(UF\)/UF\ Team\ Folder/Bivens-burn/
```

